---
title: "Welcome and ***R***edaction"
date: "useR group meeting: June 1, 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introductions

  - Name
  - Agency & Team
  - What kind of work/projects do you do? 
    - e.g. how do you use R
  - Experience with R
    - why did you learn (do you want to learn) R?
    - what is a fun thing you have used R for?
  
## What is DPI's useR group?

Semi-formal, cross agency group of primarily research analyst type folks at 
*all* levels of programming/R use/familiarity with statistics.
  
  - ~50% presentations on R-related (or project related) things that may be 
     of use/interest to the group and
     
  - ~50% informal (possibly wide ranging) discussions ... sometimes it is 
    just nice to complain at the end of a long week
    
Currently meet every 4 weeks
  
## Recent topics 

  - DPI ggplot2 theme
  - How to (productively) ask for help
  - YRBS Survey results*
  
## Things to consider in this brave new world

**Logistics**

  - Communications 
    - Listerv
    - G+ community?
    
  - Access to materials 
    - Access to internal website
    
  - Meeting Frequency
    - Cross-agency meeting every 8 weeks?
    - Rotate cross-agency meeting?
    
**Planning and Organization**

  - Topic ideas

## Redaction of Aggregate/Tabular Data

Context: For DPI, this comes up via requests for custom tables.

**Primary Suppression**

  - protecting cells that would consitute a disclosure
  - typically, there is a minimum threshold e.g. cell size of 5
  
**Secondary Suppression**

  - ensuring that primary suppressed cells are protected from indirect 
    disclosure, i.e. via subtraction
    
## Comparison of techniques for treating tabular data 

Table: (Excerpt from Australian Bureau of Statistics)

--------------------------------------------------------------------------------------------
Strategy           Advantages                       Disadvantages
------------------ -------------------------------- ----------------------------------------
Data Reduction     - Relatively easy to implement   - Does not reliably protect individuals 
                   <br>- Requires minimal education   from differencing between multiple 
                   of users                           overlapping tables <br>
                                                    - May reduce data's usefulness <br>
                                                    - The data custodian chooses what data 
                                                      to remove without necessarily knowing 
                                                      what is most important to the data 
                                                      users <br>
                                                    - Requires secondary suppression, some
                                                      suppressed ceslls may still be 
                                                      estimated
                                                    
Data Modification  - Generally does not affect the  - Does not provide additivity within 
                     data's overall utility <br>      tables unless secondary modifications
                   - Generally protects against       are applied <br>
                     differencing, zero cells,      - Requires some education of users <br>
                     and 100% cells <br>            - Entails significant setup costs <br>
                   - May be automated, requiring    - May reduce the data's usefulness 
                     minimal human input              (particularly when analyzing small
                                                      areas/populations)
--------------------------------------------------------------------------------------------

## Writing a custom function for secondary redaction

Turns out, it is hard to write your own functions for secondary suppression
even for simple tables ... let along difficult tables like those that people 
are asking for.

Enter the `sdcTable` package (*note: use the development version from github,
the version on CRAN has a bug in it*)

```{r, eval = FALSE, echo = TRUE}
devtools::install_github("bernhard-da/sdcTable")
```

If you are interested, I suggest you follow along with the vignette.

It can take a while to find the optimal solution. One run on a decently big
table took 4 hours on my desktop computer.

## Redaction Resources

 - DPI <https://dpi.wi.gov/wise/data-privacy/safeguarding>
 - `sdcTable` Vignette <https://cran.r-project.org/web/packages/sdcTable/vignettes/sdcTable.pdf>
 - US Dept of Education <https://nces.ed.gov/pubs2011/2011603.pdf>
 - ESS Handbook for Statistical Disclosure Control <https://wiki.helsinki.fi/download/attachments/51451064/SDC_Handbook.pdf> 
 - Austrailan Bureau of Statistics (Aug 2017) <http://www.abs.gov.au/AUSSTATS/abs@.nsf/ProductsbyReleaseDate/13B872F17A7DD0F8CA258184001896D6?OpenDocument>
 
Any other tools/advice


